---
title: "Chapter 13 Practice"
author: "Owen Webb"
date: today
toc: true
toc-depth: 2
format: html
editor: visual
---

# Loading Required Packages

```{r}
#| warning: false

library(dplyr)
library(dslabs)
```

# Task 1

```{r}
data(stars)
?stars
colnames(stars)
```

🎯 Using summarize() and group_by(), create a data table containing the mean values of the "temp" and "magnitude" variables in the "stars" dataset for each type of star.

```{r}
#⚡Code 
twinkle_twinkle <- stars %>% 
  group_by(type) %>% 
  summarize(
    mean_temperature = mean(temp),
    mean_magnitude = mean(magnitude),
    .groups = "drop"
  )
```

```{r}
# ✅  Checks
twinkle_twinkle
```


🗣 I am starting to see the utility in the group_by() and summarize() functions, but I don't feel like I've quite got the hang of using them. I think the part that I'm still a little confused about is the .groups = drop and what that does.

# Task 2

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅ Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.

# Task 3

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅ Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.

# Task 4

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅ Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.
