---
title: "Chapter 13 Practice"
author: "Owen Webb"
date: today
toc: true
toc-depth: 2
format: html
editor: visual
---

# Loading Required Packages

```{r}
#| warning: false

library(dplyr)
library(dslabs)
```

# Task 1

```{r}
data(stars)
?stars
colnames(stars)
```

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅  Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.

# Task 2

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅ Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.

# Task 3

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅ Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.

# Task 4

🎯 What is the goal of the code? What am I trying to learn, check, or demonstrate?

```{r}
#⚡Code 

```

```{r}
# ✅ Checks

```


🗣/🪞 What I learned / Next step. 1–2 sentences.
