---
title: "Chapter 12 Practice"
author: "Owen Webb"
date: today
toc: true
toc-depth: 2
number-sections: true
format: html
editor: visual
---

## Loading Packages

I need packages dplyr and dslabs for this practice space.

```{r}
#| warning: false

library(dplyr)
library(dslabs)
```

## Task 1: Concept Check

Vectorized conditional functions like ifelse() and case_when() will return a new vector with a result for every condition.

## Task 2: Simple ifelse()

🎯 Goal: I am trying to create a vector that accompanies the heights in the dataset "heights" that labels them as either "Tall" or "Not Tall".

```{r}
#Precheck
?dslabs::heights
heights <- dslabs::heights
names(heights)
```

🗣 The "height" column (column 2) contains the height values in this dataset. The "Tall" label should be assigned to the yes argument, since the logical test is testing for values higher than 68 inches.

```{r}
# ⚡  Code
height_label <- ifelse(test = heights$height > 68, yes = "Tall", no = "Not Tall")
```

```{r}
# ✅  Check
table(height_label)
```

🗣 The ifelse() function seemed to have worked as expected. Both the labels "Tall" and "Not Tall" show up in the table output. There are 508 Not tall and 542 Tall heights.

## Task 3: Recoding with mutate() and case_when()

🎯 Goal: Use case_when() to demonstrate the utility of the function when labeling columns with more than two categories.

```{r}
# ⚡  Code
mpg_recode <- mtcars %>% 
  mutate(mpg_class = case_when(
    mpg > 25 ~ "High MPG",
    mpg < 15 ~ "Low MPG",
    TRUE     ~ "Medium MPG"
  ))
```

```{r}
# ✅  Checks
table(mpg_recode$mpg_class)
```

🗣 The code seemed to have worked. I see the three classes I was expecting: High MPG, Low MPG, and Medium MPG. There are 6 High MPGs, 5 Low MPGs, and 21 Medium MPGs.

## Task 4: Recode - Name Focus

```{r}
# precheck
?dslabs::stars
stars <- dslabs::stars

View(stars)
colnames(stars)
summary(stars$temp)
```

🗣 The column "temp" contains the temperature values in the data frame. The name of the new column will go left of the = sign in the mutate function.

🎯 Goal: Produce a new data frame (stars_recode) with a new column called temp_group, categorizing the star temperatures as either "Cool" (temp \< 4000), "Warm" (4000 \< temp \< 7500), and "Hot" (temp \> 7500) using mutate() and case_when().

```{r}
# ⚡  Code
stars_recode <- stars %>% 
  mutate(
    temp_group = case_when(
    temp < 4000 ~ "Cool",
    temp <= 7500 ~ "Warm",
    TRUE        ~ "Hot"
   )
  )
```

```{r}
# ✅ Checks
table(stars_recode$temp_group)
```

🗣 The code worked. There are 3 categories: "Cool" with 35 obs, "Hot" with 38 obs, and "Warm" with 23 obs. Its interesting to me that the "Cool" are labeled that way even though they are also less than or equal to 7500.

## Task 5: Unit Conversion with mutate()

```{r}
#precheck
heights <- dslabs::heights
head(heights)

```

🎯 Goal: Use mutate to create a new column, converting inches into cm.

```{r}
#⚡ Code
international_friendly_heights <- heights %>% 
  mutate(height_cm = height * 2.54)

```

```{r}
#✅ Checks
head(international_friendly_heights
     )
```

🗣 It appears that the code was successful. The height values in inches were successfully all converted to centimeters. I think this is a very interesting application of mutate() and I can see lots of possible uses in this way.
