---
title: "Chapter 13 Practice"
author: "Owen Webb"
date: today
toc: true
toc-depth: 2
format: html
editor: visual
---

# Loading Required Packages

```{r}
#| warning: false

library(dplyr)
library(dslabs)
```

# Task 1

```{r}
data(stars)
?stars
colnames(stars)
```

🎯 Using summarize() and group_by(), create a data table containing the mean values of the "temp" and "magnitude" variables in the "stars" dataset for each type of star.

```{r}
#⚡Code 
twinkle_twinkle <- stars %>% 
  group_by(type) %>% 
  summarize(
    mean_temperature = mean(temp),
    mean_magnitude = mean(magnitude),
    .groups = "drop"
  )
```

```{r}
# ✅  Checks
twinkle_twinkle
```


🗣 I am starting to see the utility in the group_by() and summarize() functions, but I don't feel like I've quite got the hang of using them. I think the part that I'm still a little confused about is the .groups = drop and what that does.

# Task 2

```{r}
# Pre-checks
?olive
data(olive)
colnames(olive)
```

🎯 I need to find the mean and standard deviation of the percent oleic acid content of olive oils from different regions.

```{r}
#⚡Code 
#fatty_acid <- olive$oleic

Olive_Garden <- olive %>% 
  group_by(region) %>% 
  summarize(
    mean_value = mean(oleic, na.rm = TRUE),
    sd_value = sd(oleic, na.rm = TRUE),
    n = n(),
  )

```

```{r}
# ✅ Checks
Olive_Garden

```


🗣 I was struggling with this at first, as the check was returning the same mean and sd for every region. After troubleshooting for a while trying different fatty acids and different functions, I used the Boise State AI tool (Claude Sonnet 4) for help, using the following prompt:

"I am trying to calculate the mean oleic acid content in olive oils grouped by region (from a dslabs dataset in R). I wrote the following code, but I keep getting the same mean and standard deviation for every region.

⚡Code 
fatty_acid <- olive$oleic

Olive_Garden <- olive %>% 
  group_by(region) %>% 
  summarize(
    mean_value = mean(oleic, na.rm = TRUE),
    sd_value = sd(oleic, na.rm = TRUE),
    n = n(),
  )
Act as a tutor; guide me to the solution to this problem. "

The LLM led me to realize that the "fatty_acid <- olive$oleic" was creating a vector which was being used to calculate the mean and standard deviation, with no grouping. I fixed it by commenting out that line and then specifying in summary() to use the "oleic" column from the olive dataset. 

# Task 3

```{r}
data("us_contagious_diseases")
?us_contagious_diseases
```

```{r}
unique(us_contagious_diseases$disease)
length(unique(us_contagious_diseases$disease))
```
🗣 There are `r length(unique(us_contagious_diseases$disease))` unique diseases within the `disease` column of the `us_contagious_diseases` dataset. They are `r unique(us_contagious_diseases$disease)"`

🎯 I need to determine the number of states included in the "us_contagious_diseases" dataset.
```{r}
# Pre-check
unique(us_contagious_diseases$state)

length(unique(us_contagious_diseases$state))
```

🗣 There are `r length(unique(us_contagious_diseases$state))` unique states within the `state` column of the `us_contagious_diseases` dataset. 

# Task 4

## Step 1 - Prediction

If there are 51 states in the dataset, and you assume every state has data for all 7 diseases, how many groups will 'group_by(state, disease)' create?

I think the number of groups create would be: B. \(51 \times 7 = 357\) groups. This makes sense, because for each state, there would need to be 7 disease groups (so 7 groups 51 times).

## Step 2 - Test

🎯 I need to test my prediction to see how many groups will be made by grouping both state and disease with the group_by() function

```{r}
#⚡Code 
categorical_counts <- us_contagious_diseases %>%
  group_by(state, disease) %>%
  summarize()
```

```{r}
# ✅ Checks
count(categorical_counts)
categorical_counts
```


🗣 My prediction was correct. There are 7 rows per state, and 51 states, so there are \(51 \times 7 = 357\) groups. 

# Task 5

🎯 Determine the 3 diseases with the highest national mean counts using group_by() and summarize().

```{r}
#⚡Code 
Infectious_Burden <- us_contagious_diseases %>% 
  group_by(disease) %>% 
  summarize(
  mean_count = mean(count, na.rm = TRUE),
  .groups = "drop"
  ) %>% 
  arrange(desc(mean_count))
```

```{r}
# ✅ Checks
head(Infectious_Burden, 3)
```


🗣 The three diseases with the highest mean counts were measles, pertussis, and mumps. 

# Task 6

The group_by() function has to come before summarize() because the summarize function creates a new data frame that will not contain the variable that you want to group by. If you pipe that data frame into group_by(), R will return an error because the variable that you place in the argument of group_by() will not exist in the new summary data frame. For example, if you summarize an entire dataset before grouping and calculate the mean of a variable, the mean of the entire variable would be calculated (roughly equivalent to typing out mean(variable)). On the other hand, if you group before summarizing, the summary stats would be calculated for each grouping rather than the entire dataset. 
