---
title: "Chapter 16 Practice"
author: "Owen Webb"
toc: true
toc-depth: 2
format: html
editor: visual
---

## Loading packages
```{r}
#| warning: FALSE
library(dplyr)
library(dslabs)
library(ggplot2)
```

# Task 1

Three steps to ggplot:
  1. Assigning a dataframe: This just tells ggplot() which dataset to use for the graph/figure.
  2. Mapping aesthetics: This step assigns variables to the x and y axes of the graph.
  3. Adding geometric objects: In this step, you tell R how you want your data to be visualized, like as a scatterplot, histogram, bar graph, etc. During this step you can also do more aesthetic mapping, which will just apply to the specific geometric object added.
  
# Task 2

🎯 I need to make a scatterplot for the `iris` dataset, assigning sepal length to the x-axis and petal length to the y-axis. The points should be colored by species. 

```{r}
# ⚡ Code
Pretty_Flowers <- iris %>% 
  ggplot(aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point(size = 3) +
  xlab("Sepal Length (cm)") +
  ylab("Petal Length (cm)")
```


```{r}
# ✅ Checks
class(Pretty_Flowers)
Pretty_Flowers
```

🗣 It appears that there is overlap in sepal length between all 3 species, but the petal lengths are more distinct between the 3 species. *Iris setosa* has much shorter petals than the other two species, while the lengths of both sepals and petals overlap between *I. versicolor* and *I. virginica*.

# Task 3

🎯 I need to create a boxplot comparing the spreads of petal lengths between the 3 *Iris* species. 

```{r}
# ⚡ Code
Petal_Spread <- iris %>% 
  ggplot(aes(x = Species, y = Petal.Length, color = Species)) +
  geom_boxplot(show.legend = FALSE)
```


```{r}
# ✅ Checks
class(Petal_Spread)
Petal_Spread
```

🗣 It appears that *Iris virginica* has the greatest spread in petal lengths. Graphing the dataset as a boxplot allows me to see the spread of the values by species easier, as well as the central values and outliers for each species. 

# Task 4

🎯 I need to create a density plot comparing the distributions of CO2 intake by plant type, making the fill partially transparent, and I need to add a title to my graph. 

```{r}
# ⚡ Code
CO2_Graph <- CO2 %>% 
  ggplot(aes(x = uptake, fill = Type)) +
  geom_density(alpha = 0.5) +
  ggtitle("CO2 Uptake By Plant Type")
```


✅ What checks can I do to confirm that the code worked as expected?

```{r}
# ✅ Checks
class(CO2_Graph)
CO2_Graph
```

🗣 The overlap in distributions between the two plant types suggests that the two plant types might not significantly differ in their CO2 uptake. 

# Task 5

🎯 I need to create a graph with two layers: a histogram showing distributions of heights, and then a (partially transparent) density curve of heights layered over the histogram. I need the histogram bars to be blue and density curve to be orange.  

```{r}
# ⚡ Code
Tall_Density <- heights %>% 
  ggplot(aes(x = height)) +
  geom_histogram(aes(y = after_stat(density)),
                 size = 1,
                 fill = "blue",
                 color = "white") +
 geom_density(color = "darkorange",
            fill = "darkorange",
            alpha = 0.3,
            size = 1) +
  xlab("Height (inches)") +
  ylab("Density")
```


```{r}
# ✅ Checks
class(Tall_Density)
Tall_Density
```

🗣 The density curve provides a better visualization of the overall shape of the distribution. It does make it easier to recognize the distribution as normal (maybe with a very slight left skew).

# Task 6
