---
title: "Chapter 16 Practice"
author: "Owen Webb"
toc: true
toc-depth: 2
format: html
editor: visual
---

## Loading packages
```{r}
#| warning: FALSE
library(dplyr)
library(dslabs)
library(ggplot2)
```

# Task 1

Three steps to ggplot:
  1. Assigning a dataframe: This just tells ggplot() which dataset to use for the graph/figure.
  2. Mapping aesthetics: This step assigns variables to the x and y axes of the graph.
  3. Adding geometric objects: In this step, you tell R how you want your data to be visualized, like as a scatterplot, histogram, bar graph, etc. During this step you can also do more aesthetic mapping, which will just apply to the specific geometric object added.
  
# Task 2

🎯 

```{r}
# ⚡ Code

```


✅ What checks can I do to confirm that the code worked as expected?

```{r}
# ✅ Checks

```

🗣 

# Task 3

🎯 

```{r}
# ⚡ Code

```


✅ What checks can I do to confirm that the code worked as expected?

```{r}
# ✅ Checks

```

🗣 
# Task 4

🎯 

```{r}
# ⚡ Code

```


✅ What checks can I do to confirm that the code worked as expected?

```{r}
# ✅ Checks

```

🗣 

# Task 5

🎯 

```{r}
# ⚡ Code

```


✅ What checks can I do to confirm that the code worked as expected?

```{r}
# ✅ Checks

```

🗣 