---
title: "Chapter 12 Practice"
author: "Owen Webb"
date: today
toc: true
toc-depth: 2
number-sections: true
format: html
editor: visual
---

## Loading Packages

I need packages dplyr and dslabs for this practice space.

```{r}
#| warning: false

library(dplyr)
library(dslabs)
```

## Task 1: Concept Check

Vectorized conditional functions like ifelse() and case_when() will return a new vector with a result for every condition.

## Task 2: Simple ifelse()

🎯 Goal: I am trying to create a vector that accompanies the heights in the dataset "heights" that labels them as either "Tall" or "Not Tall".

```{r}
#Precheck
?dslabs::heights
heights <- dslabs::heights
names(heights)
```

🗣 The "height" column (column 2) contains the height values in this dataset. The "Tall" label should be assigned to the yes argument, since the logical test is testing for values higher than 68 inches.

```{r}
# ⚡  Code
height_label <- ifelse(test = heights$height > 68, yes = "Tall", no = "Not Tall")
```

```{r}
# ✅  Check
table(height_label)
```

🗣 The ifelse() function seemed to have worked as expected. Both the labels "Tall" and "Not Tall" show up in the table output. There are 508 Not tall and 542 Tall heights.

## Task 3: Recoding with mutate() and case_when()
