---
title: "Wildfire Signature Assignment"
author: "Owen Webb"
date: today
toc: true
toc-depth: 3
format: html
editor: visual
---

# Part 1

## Part A: Data Wrangling

Load packages
```{r}
#| warning: FALSE
library(readr)
library(tidyverse)
```

### Task 1

🎯 I need to import the whp_raw_sample.csv into into the workspace.

```{r}
#⚡ Code:
county_raw <- read_csv("whp_raw_sample.csv")
```
```{r}
# ✅  Explore Data
head(county_raw)
colnames(county_raw)
```

🗣 The `county_raw` dataset is made up of individual raster cells making up the map of Idaho, each cell with its own wildfire risk value (whp). The whp column seems to be a key variable for this analysis. 

### Task 2 

🎯 I need to use summarize() with county_raw to determine the number of rows in the dataset, number of missing whp values, and the min and max values in the cell_area_km2 column.

```{r}
#⚡ Code:
county_summary <- county_raw %>% 
  summarize(
    missing_whp = sum(is.na(whp) == TRUE),
    min_area = min(cell_area_km2),
    max_area = max(cell_area_km2),
    rows = n()
  )
```

```{r}
# ✅  Check
print(county_summary)
```

🗣 These numbers show that this dataset is very high quality. It appears there are no missing whp values, theres a very large sample size, and the areas of the cells don't seem to vary much at all.

### Task 3

🎯 

```{r}
#⚡ Code:

```

```{r}
# ✅  Check

```

🗣 

### Task 4

🎯 

```{r}
#⚡ Code:

```

```{r}
# ✅  Check

```

🗣 
