---
title: "Wildfire Signature Assignment"
author: "Owen Webb"
date: today
toc: true
toc-depth: 3
format: html
editor: visual
---

# Part 1

## Part A: Data Wrangling

Load packages
```{r}
#| warning: FALSE
library(readr)
library(tidyverse)
```

### Task 1

🎯 I need to import the whp_raw_sample.csv into into the workspace.

```{r}
#⚡ Code:
county_raw <- read_csv("whp_raw_sample.csv")
```
```{r}
# ✅  Explore Data
head(county_raw)
colnames(county_raw)
```

🗣 The `county_raw` dataset is made up of individual raster cells making up the map of Idaho, each cell with its own wildfire risk value (whp). The whp column seems to be a key variable for this analysis. 

### Task 2 

🎯 I need to use summarize() with county_raw to determine the number of rows in the dataset, number of missing whp values, and the min and max values in the cell_area_km2 column.

```{r}
#⚡ Code:
county_summary <- county_raw %>% 
  summarize(
    missing_whp = sum(is.na(whp) == TRUE),
    min_area = min(cell_area_km2),
    max_area = max(cell_area_km2),
    rows = n()
  )
```

```{r}
# ✅  Check
print(county_summary)
```

🗣 These numbers show that this dataset is very high quality. It appears there are no missing whp values, theres a very large sample size, and the areas of the cells don't seem to vary much at all.

### Task 3

🎯 I need to use quantile() to compute the quantiles of the whp column

```{r}
#⚡ Code:
e <- quantile(county_raw$whp, probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1), na.rm = TRUE)

```

```{r}
# ✅  Check
print(e)
```

🗣 Object e represents the different percentiles of the whp variable. This is helpful because it gives me a rough idea of the distribution of the values, and what values are more extreme or more central.

### Task 4

🎯 I need to categorize rows in the county_raw dataset by their relative wildfire risk.

```{r}
#⚡ Code:
county_cats <- county_raw %>% 
  mutate(Category = case_when(
    whp >= 0 & whp < 130.7305 ~ "Very Low",
    whp >= 130.7305 & whp < 397.6562 ~ "Low",
    whp >= 397.6562 & whp < 1101.0156 ~ "Moderate",
    whp >= 1101.0156 & whp < 3479.8125 ~ "High",
    whp >= 3479.8125 & whp < 7708.5383 ~ "Very High",
    whp >= 7708.5383 & whp < 22046.7472 ~ "Extreme",
    whp >= 22046.7472 & whp <= 63465.4219 ~ "Ultra (Top 1%)"
  ))
county_cats$Category <- factor(county_cats$Category, levels = c("Very Low", "Low", "Moderate", "High", "Very High", "Extreme", "Ultra (Top 1%)"), ordered = TRUE)
```

```{r}
# ✅  Check
county_cats %>% group_by(Category) %>% count()
glimpse(county_cats)
```

🗣 I initially tried formatting my case_when arguments as `x <= whp < y`, but that didn't work, so I changed my formatting to use the & operator instead and that worked. 

It seems as though, luckily, the majority of Idaho's land is at Very Low to Moderate fire risk, but there's still a troubling amount (~25%) that is at High to Ultra fire risk. 
