---
title: "Wildfire Signature Assignment"
author: "Owen Webb"
date: today
toc: true
toc-depth: 3
format: html
editor: visual
---

## Part A: Data Wrangling

Load packages
```{r}
#| warning: FALSE
library(readr)
library(tidyverse)
```

### Task 1

🎯 I need to import the whp_raw_sample.csv into into the workspace.

```{r}
#⚡ Code:
county_raw <- read_csv("whp_raw_sample.csv")
```
```{r}
# ✅  Explore Data
head(county_raw)
colnames(county_raw)
```

🗣 The `county_raw` dataset is made up of individual raster cells making up the map of Idaho, each cell with its own wildfire risk value (whp). The whp column seems to be a key variable for this analysis. 

### Task 2 

🎯 I need to use summarize() with county_raw to determine the number of rows in the dataset, number of missing whp values, and the min and max values in the cell_area_km2 column.

```{r}
#⚡ Code:
county_summary <- county_raw %>% 
  summarize(
    missing_whp = sum(is.na(whp) == TRUE),
    min_area = min(cell_area_km2),
    max_area = max(cell_area_km2),
    rows = n()
  )
```

```{r}
# ✅  Check
print(county_summary)
```

🗣 These numbers show that this dataset is very high quality. It appears there are no missing whp values, theres a very large sample size, and the areas of the cells don't seem to vary much at all.

### Task 3

🎯 I need to use quantile() to compute the quantiles of the whp column

```{r}
#⚡ Code:
e <- quantile(county_raw$whp, probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1), na.rm = TRUE)

```

```{r}
# ✅  Check
print(e)
```

🗣 Object e represents the different percentiles of the whp variable. This is helpful because it gives me a rough idea of the distribution of the values, and what values are more extreme or more central.

### Task 4

🎯 I need to categorize rows in the county_raw dataset by their relative wildfire risk.

```{r}
#⚡ Code:
county_cats <- county_raw %>% 
  mutate(Category = case_when(
    whp >= 0 & whp < 130.7305 ~ "Very Low",
    whp >= 130.7305 & whp < 397.6562 ~ "Low",
    whp >= 397.6562 & whp < 1101.0156 ~ "Moderate",
    whp >= 1101.0156 & whp < 3479.8125 ~ "High",
    whp >= 3479.8125 & whp < 7708.5383 ~ "Very High",
    whp >= 7708.5383 & whp < 22046.7472 ~ "Extreme",
    whp >= 22046.7472 & whp <= 63465.4219 ~ "Ultra (Top 1%)"
  ))
county_cats$Category <- factor(county_cats$Category, levels = c("Very Low", "Low", "Moderate", "High", "Very High", "Extreme", "Ultra (Top 1%)"), ordered = TRUE)
```

```{r}
# ✅  Check
county_cats %>% group_by(Category) %>% count()
glimpse(county_cats)
```

🗣 I initially tried formatting my case_when arguments as `x <= whp < y`, but that didn't work, so I changed my formatting to use the & operator instead and that worked. 

It seems as though the majority of Idaho's land is at Very Low to Moderate fire risk, but there's still a troubling number of cells that are at High to Ultra fire risk. 

###  Task 5

🎯 I need to create a summary table that has 4 variables: Category, area_km2 (sum of the areas of all the cells), n_cells (number of grid cells), and Probability (area_km2 divided by sum of area_km2)

```{r}
#⚡ Code:
prob_tbl <- county_cats %>% 
  group_by(Category) %>% 
  summarize(
    area_km2 = sum(cell_area_km2),
    n_cells  = n(),
    .groups = "drop") %>% 
  mutate(
    Probability = area_km2/sum(area_km2)) %>% 
  select(Category, area_km2, n_cells, Probability)


```

```{r}
# ✅  Check
print(prob_tbl)
sum(prob_tbl$Probability)
```

🗣 It appears that approximately 25% of Idaho's land is at at least High risk of wildfire. The probability represents the chance of randomly selecting a cell that is in a given category. 

## Part B: Probability Modeling

### Task 1

🎯 Using a lognormal distribution, I need to calculate the probability that a fire will spread 10 or more kilometers after 8 hours in a "High" risk zone.

```{r}
#⚡ Code:
p_high_exact <- plnorm(10, meanlog = 1.907, sdlog = 0.280, lower.tail = FALSE)
```
```{r}
# ✅  Check
print(p_high_exact)
p_high_exact*100
```

🗣 Using a lognormal model, it appears that a fire has a 7.886% chance of burning 10 km or more after 8 hours. 

### Task 2

🎯 Using qlnorm(), I need to determine the 8-hour run distance that only the top 5% of fires exceed. 

```{r}
#⚡ Code:
q95_high <- qlnorm(0.95, meanlog = 1.907, sdlog = 0.280)
```
```{r}
# ✅  Check
print(q95_high)
```

🗣 95% of fires in High risk zones burn less than 10.67131 km in an 8 hour run. 

### Task 3

🎯 Using the High Risk lognormal model, I need to compare the likelihood of a fire burning 3 km vs 10 km after 8 hours.

```{r}
#⚡ Code:
d_3 <- dlnorm(3, meanlog = 1.907, sdlog = 0.280)
d_10 <- dlnorm(10, meanlog = 1.907, sdlog = 0.280)
```
```{r}
# ✅  Check
print(d_3)
print(d_10)
```

🗣 Based on the densities at each point (3km and 10km), it seems that a 10 km distance is more locally likely as the distribution has a higher density at X = 10 km.

### Task 4

#### Task 4.1

🎯 I need to simulate 450 8-hour fire distances in the High risk category.

```{r}
#⚡ Code:
set.seed(2025)
High_Lognorm <- rlnorm(450, meanlog = 1.907, sdlog = .280)
```
```{r}
# ✅  Check
summary(High_Lognorm)
length(High_Lognorm)
```

🗣 The range of values in `High_Lognorm` is approximately 12 km, and the mean is 6.987, close to the mean in the 8-hour distance table provided in the instructions. Based solely on the summary, it seems that the distribution isn't skewed much, as the mean and median are close, and they sit in the approximate middle of the min and max values. 

#### Task 4.2

🎯 

```{r}
#⚡ Code:

```
```{r}
# ✅  Check

```

🗣 

#### Task 4.3

🎯 

```{r}
#⚡ Code:

```
```{r}
# ✅  Check

```

🗣 
